The Impact of the Severe Weather Events on Human Health and Economy
======================
## Synopsys

The goal of this study was to find out the impact of the most devastating weather events both on the population health and economy. The basis of the study is the information of the NOAA Storm Database representing observations which was made across the United States in the period from 1950 to 2011 year.  
To achieve this goals we have made processing of the data represented by National Weather Service Storm Data Documentation.    
Brief straightforward analysis of given graphs shows that most devastating weather event in the USA is tornadoes followed by floods and heat.  


## Source Data 

Please download file from  
* Storm Data [47Mb] (https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)

## Data Processing
Loading libraries
```{r results='hide', messages=FALSE, warning=FALSE}
library(data.table)
require(R.utils)
library(plyr)
library(ggplot2)
library(grid)
library(gridExtra)
```
Unzipping bz2 and reading data from .csv file
```{r results='hide'}
zip_file <- 'repdata-data-StormData.csv.bz2'
file <- 'repdata-data-StormData.csv'
if (!file.exists(file) & file.exists(zip_file) ){
        bunzip2(zip_file, remove = FALSE)
} else if(!file.exists(file) & !file.exists(zip_file)){
        print ('Please download https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2 file to source directory')
}
data <- fread(file)
```
Calculating totals by the type of the events (victims for injuries/fatalities, million dollars for economy damage)
```{r}
data$EVTYPE <- factor(data$EVTYPE)
totals <- ddply(data,.(EVTYPE), summarize, injuries=sum(INJURIES), fatalities=sum(FATALITIES), propdmg=sum(PROPDMG)/1000000, cropdmg=sum(CROPDMG)/1000000)
```
Subsetting events with significant values
```{r}
tot1<-totals[(totals$injuries>5000 | totals$fatalities>400),] # Filtering the most harmful events
tot2<-totals[(totals$propdmg>0.5 | totals$cropdmg>0.05),] # Filtering the most devastating events
```

## Results
### Most harmful events affecting on on the population
```{r}
p1 <- ggplot(tot1, aes(x=EVTYPE, y=injuries, fill=EVTYPE)) + geom_bar(stat="identity") + theme(axis.ticks = element_blank(), axis.text.x = element_blank(), axis.title.x = element_text(size=11), axis.title.y = element_text(size=11)) + xlab('Event Type')
p2 <- ggplot(tot1, aes(x=EVTYPE, y=fatalities, fill=EVTYPE)) + geom_bar(stat="identity") + theme(axis.ticks = element_blank(), axis.text.x = element_blank(), axis.title.x = element_text(size=11), axis.title.y = element_text(size=11)) + xlab('Event Type')
g1<-grid.arrange(p1, p2, nrow=2, main="Figure 1. Most harmful events. Population")
```
  
The analysis shows that most harmful for human health are Tornadoes and Heat. 

### Most harmful events affecting on on the economy 
```{r}
q1 <- ggplot(tot2, aes(x=EVTYPE, y=propdmg, fill=EVTYPE)) + geom_bar(stat="identity") + theme(axis.ticks = element_blank(), axis.text.x = element_blank(), axis.title.x = element_text(size=11), axis.title.y = element_text(size=11)) + xlab('Event Type') + ylab('Property Damage, Mil USD')
q2 <- ggplot(tot2, aes(x=EVTYPE, y=cropdmg, fill=EVTYPE)) + geom_bar(stat="identity") + theme(axis.ticks = element_blank(), axis.text.x = element_blank(), axis.title.x = element_text(size=11), axis.title.y = element_text(size=11)) + xlab('Event Type') + ylab('Crop Damage, Mil USD')
g2<-grid.arrange(q1, q2, nrow=2, main="Figure 2. Most harmful events. Economy")
```
  
The analysis shows that most harmful for economy are Tornadoes and Floods(different types).   
Hails have significanat impact on the crop.
